---
title: "Kruskal"
author: "Damir Pintar"
date: "October 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(igraph)
library(arules)
library(optrees)
library(ggplot2)
library(ggraph)
library(stringr)
library(gridExtra)


# change the working directory according to your custom file structure
setwd("~/R/Kruskal/RMD")
source("../R/helperFunctions.R")
```

## Input dataset loading and pre-processing

We will use two datasets - a toy dataset to showcase the functions and methodology, and a real-life education dataset for actual analysis.

```{r dfs_init}
toy.df <-data.frame( A = c(T, T, T, F, F, F, T, T, T, T, F, F),
                     B = c(T, T, T, F, F, F, F, T, T, T, F, F),
                     C = c(F, F, F, T, T, T, F, T, F, T, F, F),
                     D = c(F, F, F, T, T, T, F, F, F, T, F, F),
                     E = c(F, F, F, F, F, F, T, T, T, T, F, F),
                     F = c(T, F, F, F, F, F, F, F, F, F, T, T))

edu.df <- read.csv("../datasets/MI2012.csv")
edu.df <- edu.df[, -which(sapply(edu.df, sum)==0)]
edu.df <- lapply(edu.df, as.logical) 
edu.df <- as.data.frame(edu.df)

edu2.df <- read.csv("../datasets/MI2010.csv")
edu2.df <- edu2.df[, -length(edu2.df)] # removing 'fake_class', a remnant from Orange
#edu2.df <- edu2.df[, -which(sapply(edu2.df, sum)==0)]
edu2.df <- lapply(edu2.df, as.logical) 
edu2.df <- as.data.frame(edu2.df)


edu3.df <- edu2.df[, grep("Q", names(edu2.df))]
edu3.df <- lapply(edu3.df, as.logical) 
edu3.df <- as.data.frame(edu3.df)

comp.df <- read.csv("../datasets/Computer_shop.csv")
comp.df <- comp.df[, -ncol(comp.df)]  #redundant column
comp.df <- lapply(comp.df, as.logical) 
comp.df <- as.data.frame(comp.df)
```


Choose the input dataset and support here:

```{r choose}
df <- edu.df
support <- 0.50
encodeNames <- T
```


## Input dataset loading and pre-processing

Let’s encode the dataset (and keep the info about the encoding in a separate data frame).

```{r dfColEnc}
dfColEncodings <- data.frame(names = names(df), code = names(df))
if (encodeNames) {
  dfColEncodings <- alphaColNames(names(df))
  names(df) <- dfColEncodings$code
}
```



## Apriori

We will mine the frequent itemsets using the *eclat* algorithm and check out the summary of the results.

```{r miningItemsets}
# we are closed mining frequent itemsets using the eclat algorithm
itemsets <- eclat(df, parameter = list(supp = support,
                                               target = "closed frequent itemsets"))


# we will mine frequent itemsets too to get the singletons
# in order to have them in the final graph
# (if they are not closed,  they would not appear in the CFI mining process)
itemsetsTEMP <- eclat(df, parameter = list(supp = support,
                                               target = "frequent itemsets"))

# now we keep just the singleton itemsets
quality(itemsetsTEMP)$noItems <- sapply(as(items(itemsetsTEMP), "list"), length)
singletonsInd <- quality(itemsetsTEMP)$noItems == 1

# and put them in the original dataset
itemsets <- union(itemsets, itemsetsTEMP[singletonsInd,])

# possible alternative: apriori (eclat is faster!)
# itemsets2 <- apriori(df, parameter = list(supp = 1/12,
#                                              target = "frequent itemsets"))

#summary of results
summary(itemsets)
```

```{r dfConv}
res <- DATAFRAME(itemsets)
res
```

Few other methods we can do with the itemset:

```{r dfConv2}
# convert the results to a list
res.list <- as(items(itemsets), "list")

# check out the default visualization
image(items(itemsets))
```


We can add more quality measures to mined itemsets:

```{r addingQualities}
# adding the size of the itemset as a parameter
quality(itemsets)$noItems <- sapply(res.list, length)
quality(itemsets)$absSup <- round(quality(itemsets)$support * nrow(df))

# let's add the full name to singleton collections
quality(itemsets)$fullName <- ""
if (encodeNames) {
  for (i in 1:nrow(quality(itemsets))) 
    if (quality(itemsets)[i, "noItems"] == 1)
      quality(itemsets)[i, "fullName"] <- as.character(dfColEncodings[dfColEncodings$code ==
                    substr(res[i, "items"], 2, nchar(as.character(res[i, "items"])) - 1), "names"])
}

inspect(itemsets)
```



## Building graphs

We will now build graphs using three different algorithms - Kruskal, TD strategy, BU strategy.

```{r Kruskal}
graphKruskal <- msTreeKruskalCustom(itemsets)
#graphKruskalRestricted <- msTreeKruskalCustom(itemsets, root = "{C,D}")
```

```{r stratTU}
graphTDstrategy <- treeTDstrategy(itemsets)
```

```{r stratBU}
graphBUstrategy <- treeBUstrategy(itemsets, verbose = T)
  
```

## Visualizing graphs

```{r}
g <- graphKruskal
layo <- create_layout(g, layout = 'tree', root = 2)
ggraph(layo) + geom_edge_arc(aes(label = weight), colour = "grey", curvature = 0.25) + 
 # geom_node_circle(aes(r = 0.5), fill = "white") + 
  geom_node_text(aes(label = name)) + 
  geom_node_text(aes(label = fullName), colour = "blue", nudge_y = -0.5) +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank()) 
```


```{r}
g <- graphKruskal
V(g)$hasFullName <- V(g)$fullName == ""
ggraph(g, layout = 'lgl') +  
  geom_edge_arc(aes(label = weight), colour = "grey", curvature = 0.25, 
                arrow = arrow(angle = 30, length = unit(0.125, "inches"), type = "closed")) +  
     geom_node_circle(aes(r = 2.5 * support, fill = as.factor(hasFullName)), size = 0.1, alpha = 0.75) + 
     scale_fill_manual(values=c("#FFFFFF", "#CCEECC"), guide=FALSE) +
  geom_node_text(aes(label = name)) +
  geom_node_text(aes(label = fullName), colour = "blue", nudge_y = -2.5) + theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank()) 
ggsave("grafo.png", height = 10, width = 14)
```




## Top-down strategy graph 

```{r}
g <- graphTDstrategy

layo <- create_layout(g, layout = 'kk')

#creating custom layout
# only for comp.df and supp 0.8
if (length(V(g)) == 11) {
  layo$x <- c(20, 10, 52, 82, 5, 15, 60, 30, 45, 75, 90)
  layo$x <- layo$x / 3
  layo$y <- c(50, 40, 40, 40, 20, 27, 20, 20, 27, 27, 20)
}

ggraph(layo) + geom_edge_arc(colour = "grey", curvature = 0.025) + 
    #geom_node_circle(aes(r = 2.5), fill = "white") + 
    geom_node_text(aes(label = name)) + 
    geom_node_text(aes(label = fullName), colour = "blue", nudge_y = -3.5) +
    theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()) 
```


```{r}
ggraph(g) +  geom_edge_arc(colour = "grey", curvature = 0.25) +  
    geom_node_circle(aes(r = 0.3), fill = "white") + geom_node_text(aes(label = name)) +
  geom_node_text(aes(label = fullName), colour = "blue", nudge_y = -0.5) + theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank()) 
```

## Bottom-up strategy graph

```{r}
g <- graphBUstrategy

ggraph(g) +  geom_edge_arc(colour = "grey", curvature = 0.25) +  
    geom_node_circle(aes(r = 0.3), fill = "white") + geom_node_text(aes(label = name)) +
  geom_node_text(aes(label = fullName), colour = "blue", nudge_y = -0.5) + theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank()) 
```